---
title: "Skeleton Connect SRA dbGaP Metadata"
output: html_notebook
---

# Load libraries
```{r}
library(RSQLite)
library(tidyverse)
library(stringr)
```

# Load in metadata and SRA sqlite
You have to set `sqlfile` to the path of the sqlite file on your computer
```{r}
load('../data/demo_dbgap_metadata.Rdata')
sqlfile <- '~/SRAmetadb.sqlite'
sra_con <- dbConnect(SQLite(), sqlfile)
```

# Left join SRAdb to dbGaP metadata
```{r}
dbGetQuery(sra_con, "select * from study where study_alias like 'phs%' limit 40")
dbGetQuery(sra_con, "select * from sra where study_alias is 'phs000257_56' limit 40")
```

# You can also store the results as a data frame
```{r}
phs000257_56 <- dbGetQuery(sra_con, "select * from sra where study_alias is 'phs000257_56'")
phs000257_56 %>% sample_n(10)
```

# Let's take a quick look at the dbGaP Metadata tables
10 random rows from each
```{r} 
tables$study_dataset_info %>% sample_n(10) %>% data.frame()
tables$study_id_variable_name %>% sample_n(10) %>% data.frame()
tables$study_info%>% sample_n(10) %>% data.frame()
tables$study_variable_code_value %>% sample_n(10) %>% data.frame()
tables$study_variable_info %>% sample_n(10) %>% data.frame()
```

# Process sra and dbGaP metadata to make join-able
```{r}
sra_in_dbGaP <- dbGetQuery(sra_con, "select * from sra where study_alias like 'phs%'")
sra_in_dbGaP <- sra_in_dbGaP %>% 
              rowwise() %>% 
              mutate(study_alias2 = str_split(study_alias, '_')[[1]][1])
tables$study_variable_info <- tables$study_variable_info %>% 
  rowwise() %>% 
  mutate(study_alias = str_split(study_accession, '\\.')[[1]][1])
```

How many studies have run level info?
```{r}
unique(tables$study_variable_info$study_alias) %in% unique(sra_in_dbGaP$study_alias2) %>% 
  table()

# Don't run unless you have a LOT of memory and CPU
#left_join(tables$study_variable_info, sra_in_dbGaP, by = c('study_alias' = 'study_alias2')) %>% group_by(study_alias) %>% summarise(Count=n()) %>% dim()
```
How many studies with emphysema information have RNA-seq data?
```{r}
rna_studies <- sra_in_dbGaP %>% 
  filter(grepl('RNA', library_strategy, ignore.case=T)) %>% 
  pull(study_alias2) %>% 
  unique()

emphysema_studies <- tables$study_variable_info %>% 
  filter(grepl('emphysema', description)) %>% 
  pull(study_alias) %>% 
  unique()

emphysema_studies %in% rna_studies %>% table()
```


# Other examples highlighting dplyr
https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html
## Studies with 'asthma' in the description
```{r}
tables$study_variable_info %>% filter(grepl("asthma", description, ignore.case=T)) %>% sample_n(10)
```

## Number of studies (and versions) in study_variable_info
```{r}
tables$study_variable_info %>% group_by(study_accession) %>% summarise(Count=n()) %>% nrow()
```

## Select three columns from study_variable_info in rows where cancer is in the description and taking 20 random rows from that
```{r}
tables$study_variable_info %>% select(study_accession, variable_accession, male_count, female_count, description) %>% filter(grepl('cancer', description, ignore.case=T)) %>% sample_n(20)
```

## Select useful columns from study_variable_info in rows where cancer is in the description and pull the studies where there are more than 50 males and 1000 females in the cohort
```{r}
interesting_studies <- tables$study_variable_info %>% 
  select(study_accession, variable_accession, male_count, female_count, description) %>%
  filter(grepl('cancer', description, ignore.case=T)) %>% 
  filter(male_count > 50, female_count > 1000) %>% 
  rowwise() %>% 
  mutate(study_accession2 = str_split(study_accession, '\\.')[[1]][1]) %>% 
  group_by(study_accession2) %>% 
  summarise(Count = n()) %>% 
  pull(study_accession2)
print('')
interesting_studies
```
